Shapes analysis for time series.

### 2024-05-13 NEURIPS24/paper.tex
up

%\paragraph*{Regularizing the problem to derive uniqueness.}
It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the space of deformations.
Following this geometric point of view, geodesics can be constructed on the space of deformations by using the norm of $\msv$. 
More precisely, on the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$,
 the following squared norm can be defined
 \begin{equation}
  \label{eq:geodesics_original}
    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
 \end{equation}
     as the minimal "energy" needed to perform the deformation $g$.
By \citep[Theorem 6]{glaunes2005transport}, there exists $v^*\in \rml^2([0,1],\msv)$ such that the previous infimum is a minimum in $v^*$ such that $(\phi^{v^*}_\tau)_{\tau\in[0,1]}$ can be understood as the geodesic between the identity function and $g$.
However, given a diffeomorphism $g$, computing $\mathscr{R}(g)$ is intractable in most cases. 
To circumvent this issue, another characterization of geodesics can be considered.
 As in Riemannian geometry, 
 instead of defining geodesics from their starting and end points, it is possible to define them from their starting point (here, the identity function) and an initial velocity $v_0\in\msv$.
 \sam{To CHANGE}
 In other words, an Exponential map on deformations is wanted:
given $v_0\in\msv$, it has been suggested to generate diffeomorphisms as $\varphi^{\{v_0\}}=\phi^v$ with
%  \begin{equation}
%   \label{eq:geodesics_shooting}
%   v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
%  \end{equation}
% since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
 By setting $\msv$ as an RKHS, the geodesic shooting problem \eqref{eq:geodesics_shooting} has a unique solution and becomes tractable, as described in the next section.
% Moreover, we have $\mathscr{R}(\varphi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in \rml^2([0,1],\msv)$.
%  Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
%  Moreover, we have $\mathscr{R}(\phi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in L^2([0,1],\msv)$.
%  %Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
%  This is presented with more details in the next paragraph.

>>>>

%\paragraph*{Regularizing the problem to derive uniqueness.}
It has been proposed in \cite{miller2006geodesic} to interpret $\msv$ as a tangent space relative to the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$.
Following this geometric point of view, geodesics can be constructed on $\msh$ by using the following squared norm 
 \begin{equation}
  \label{eq:geodesics_original}
    \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
 \end{equation}
By deriving differential constraints related to the minimum of \eqref{eq:geodesics_original} and using Cauchy lipshcitz conditions, geodesics can be defined only by giving the starting point and the initial velocity $v_0\in \msv$, as straight lines in Euclidean space.
Denoting by $w(v_0)$ the geodesic starting from the identity with inital velocity $v_0$, the exponential map is defined as $\varphi^{\{v_0\}}\triangleq \phi^v$ and the previous matching problem becomes a \textit{geodesic shooting problem}:
 \begin{equation}
  \label{eq:geodesics_shooting}
  \inf_{v_0 \in \msv} \mathscr{L}(\varphi^{\{v_0\}}.\mathbf{x},\mathbf{y}).
 \end{equation}
 Using $\varphi^{\{v_0\}}$ instead of $\phi^v$ for any $v\in \rml^2([0,1],\msv)$ regularizes the problem and induces a sparse representation for the learning diffeomorphisms.
 Moreover, by setting $\msv$ as an RKHS, the geodesic shooting problem has a unique solution and becomes tractable, as described in the next section.

% More precisely, on the group of diffeomorphisms $\msh=\{ \phi^v:\eqsp v\in \rml^2([0,1],\msv)\}$,
%  the following squared norm can be defined
%  \begin{equation}
%   \label{eq:geodesics_original}
%     \mathscr{R}^2: g\in \msh\mapsto \inf_{ v\in \rml^2([0,1],\msv):\eqsp g=\phi^v} \int_0^1 || v_\tau||_\msv^2\dd \tau
%  \end{equation}
%      as the minimal "energy" needed to perform the deformation $g$.
% By \citep[Theorem 6]{glaunes2005transport}, there exists $v^*\in \rml^2([0,1],\msv)$ such that the previous infimum is a minimum in $v^*$ such that $(\phi^{v^*}_\tau)_{\tau\in[0,1]}$ can be understood as the geodesic between the identity function and $g$.
% However, given a diffeomorphism $g$, computing $\mathscr{R}(g)$ is intractable in most cases. 
% To circumvent this issue, another characterization of geodesics can be considered.
%  As in Riemannian geometry, 
%  instead of defining geodesics from their starting and end points, it is possible to define them from their starting point (here, the identity function) and an initial velocity $v_0\in\msv$.
%  \sam{To CHANGE}
%  %In other words, an Exponential map on deformations is wanted:
% given $v_0\in\msv$, it has been suggested to generate diffeomorphisms as $\varphi^{\{v_0\}}=\phi^v$ with
% %  \begin{equation}
% %   \label{eq:geodesics_shooting}
% %   v=\underset{w\in \rml^2([0,1],\msv):\eqsp w_0=v_0}{\argmin} \int_0^1 || w_\tau||_\msv^2\dd \tau\eqsp ,
% %  \end{equation}
% % since with this definition it holds $\mathscr{R}^2(\phi^v)= \int_0^1 || v_\tau||_\msv^2\dd \tau $.
%  By setting $\msv$ as an RKHS, the geodesic shooting problem \eqref{eq:geodesics_shooting} has a unique solution and becomes tractable, as described in the next section.
% Moreover, we have $\mathscr{R}(\varphi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in \rml^2([0,1],\msv)$.
%  Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
%  Moreover, we have $\mathscr{R}(\phi^{\{v_0\}})=||v_0||_\msv$ such that we can work with $v_0\in \msv $ instead of $v\in L^2([0,1],\msv)$.
%  %Therefore, compared to CNF, only $v_0$ is parametrized by using the geodesic structure and not each $v_\tau$ for any $\tau\in[0,1]$.
%  This is presented with more details in the next paragraph.
